Towards Comparable Active Learning

### 2023-06-02 doc/al_benchmark_paper.tex

\begin{abstract}
	%Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
	In this paper we address the issue of inconsistent results in active learning (AL) literature.
	Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results. This inconsistency leads to a chaotic landscape of AL algorithms.
	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains; Tabular, Image and Text.
	Furthermore, we discuss overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
	We propose a framework for choosing hyperparameters of AL setups and accurately controlling the experiments.
	Using this framework, we report empirical results for 6 algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
{\color{red} Version: Braindump}\\
\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_bald}
	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
	\label{fig:ablation_bald}
\end{wrapfigure}
Many different algorithms have been proposed for active learning. 
In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
Uncertainty-based approaches use metrics to measure the classifiers state.
Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
%
Some previous work also aimed to provide a benchmark suite for active learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
\end{equation}
where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
A good AL algorithm should be able to perform well in both scenarios. 
Since AL has high variance and is prone to outliers, we propose to report median AUC instead of mean AUC values. \\ [1mm]
The AUC is dependent on the chosen budget, so we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Oracle}\label{sec:oracle}
Posing active learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure would be the set that induces the highest AUC score for a given budget.
However, since this problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence.
\cite{zhou2021towards} has used simulated annealing to search for the optimal sequence and used the best  solution found after a fixed time budget. 
Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
\begin{algorithm}
	\caption{Oracle}\label{alg:oracle}
	\begin{algorithmic}[1]
		\Require $\mathcal{U}, \mathcal{L}, \mathcal{Y}, \mathcal{D}_\text{test}$ Train, Margin, $\tau, \hat y_\theta$ 
		\State $\text{acc} \gets \operatorname{Train}(\mathcal{L}, \mathcal{D}_\text{test}, \hat y_\theta)$ 
		%\State $u \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}|)$
		\State $r^* \gets 0$
		\For{$t := 1 \ldots \tau$}
		\State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(\mathcal{U}_t, y_t)\}$
		\State $\text{acc}' \gets \operatorname{Train}(\mathcal{L}', \mathcal{D}_\text{test}, \hat y_\theta)$  
		\State $r \gets \text{acc} - \text{acc}'$
		\If{$r > r^*$} 
		\State $r^* \gets r$
		\State $u^* \gets \mathcal{U}_t$
		\EndIf
		\EndFor
		\If{$r^* = 0$}
		\State $u^* \gets \operatorname{margin}(\mathcal{U}, \hat y_\theta)$
		\EndIf
		\Return $u^*$
	\end{algorithmic}
\end{algorithm}
In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
Our oracle simply evaluated every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting test performance.
The point with the best test performance is selected and added to the labeled pool for that iteration.
We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
To circumvent this problem, we introduced margin sampling as a fallback option for the oracle.
Whenever the oracle does not find an unlabeled point that results in an increase in performance (indicating an overfitting position), it defaults to margin sampling in that iteration.
The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
In the algorithm $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
Alg. \ref{alg:oracle} replaces the algorithm in the AL process.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Available Information}
At each iteration $i$ the AL algorithm needs to pick an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \theta^{(i)}, \text{opt}_\theta\}$, where $\theta^{(i)}$ is the current classifier and $\text{opt}_\theta$ is the optimizer used to fit $\theta^{(i)}$.
We allow algorithms to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for the classifier, or the test/validation set.

@@

%\begin{table}[]
%	\centering
%	\begin{tabular}{l|lll}
%		& Splice        & DNA           & USPS          \\
%		\hline
%		Oracle          & 0.830 +- 0.01 & 0.836 +- 0.02 & 0.823 +- 0.01 \\
%		SAL & 0.799 +- 0.01 & 0.797 +- 0.03 & 0.809 +- 0.01 \\
%		Coreset & 0.800 +- 0.01 & 0.795 +- 0.03 & 0.787 +- 0.02 \\
%		TypiClust       & 0.790 +- 0.01 & 0.771 +- 0.04 & 0.761 +- 0.02 \\
%		MarginScore     & 0.797 +- 0.02 & 0.795 +- 0.04 & 0.808 +- 0.01 \\
%		ShannonEntropy  & 0.799 +- 0.02 & 0.794 +- 0.04 & 0.807 +- 0.01 \\
%		RandomAgent     & 0.788 +- 0.01 & 0.765 +- 0.03 & 0.772 +- 0.01 \\
%		Badge           & 0.807 +- 0.01 & 0.769 +- 0.06 & 0.797 +- 0.02 \\
%		BALD            & 0.811 +- 0.01 & 0.743 +- 0.04 & 0.717 +- 0.05
%	\end{tabular}
%	\caption{Median AUC values and standard deviation for all algorithms on the tabular datasets. Higher is better.}
%\end{table}
%
\begin{figure}
	\centering
	\includegraphics[width=0.49\linewidth]{img/eval_splice}
	\hspace{1mm}
	\includegraphics[width=0.49\linewidth]{img/eval_dna}\\[2mm]
	\includegraphics[width=0.49\linewidth]{img/micro_splice}
	\hspace{1mm}
	\includegraphics[width=0.49\linewidth]{img/micro_dna}
	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
	\label{fig:eval_vector}
\end{figure}
%
\begin{figure}
\centering
	\includegraphics[width=0.49\linewidth]{img/macro_vector}
	\includegraphics[width=0.49\linewidth]{img/macro_vector_enc}
	\includegraphics[width=0.49\linewidth]{img/macro_img}
	\includegraphics[width=0.49\linewidth]{img/macro_img_enc}
	\includegraphics[width=0.49\linewidth]{img/macro_text}
	\includegraphics[width=0.49\linewidth]{img/all_data.png}
	\caption{Critical Difference Diagram for all algorithms grouped by domain and all domains combined. Ranks are computed based on median AUC for each algorithm and dataset combination. Lower ranks are better.}
	\label{fig:cd_diagrams}
\end{figure}
%
From Fig. \ref{fig:eval_vector} we notice drastically different qualities for the same AL algorithm for different datasets.
We would like to highlight that both datasets are tabular from the medical domain with similar number of features and classes, yet we see that i.e. BALD is the best algorithm for Splice and the worst algorithm for DNA.
These inconsistencies are present between the datasets of all our tested domains, further highlighting the difficulties for comparing AL algorithms in terms of average performance.
In order to provide a meaningful analysis of which algorithm can be expected to perform best on average we ranked the algorithm for each dataset based on their median AUC and displayed these rankings in critical difference diagrams \cite{IsmailFawaz2018deep}.
In Fig. \ref{fig:cd_diagrams} we report the rankings split by domain as well as across all domains (excluding the toydata). \\ [2mm]
%
BALD performs bad with linear classifiers since they are trained without dropout and cannot cope well with missing inputs. \\
TypiClust is better with embedded data not only due to lower budgets. On other datasets it is not able to outperform other algorithms in early stages

\section{Conclusion}

@@

\section{Alternative Plot for Restarts Ablation}
\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts_2.png}
\end{figure}

>>>>

\begin{abstract}
	Active Learning has received significant attention in the field of machine learning for its potential in selecting the most informative samples for labeling, thereby reducing data annotation costs. However, the lack of reproducibility of performance gains reported in recent literature has created a chaotic landscape in Active Learning research. 
	This paper addresses these issues of inconsistent results in active learning (AL) literature.
	%Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results. This inconsistency leads to a chaotic landscape of AL algorithms.
	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains: Tabular, Image and Text.
	Furthermore, we highlight overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
	To tackle these challenges, we propose an experimental protocol to accurately control the experiments.
	We report empirical results for 6 widely used algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_bald}
	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
	\label{fig:ablation_bald}
\end{wrapfigure}
Many different algorithms have been proposed for active learning. 
In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
Uncertainty-based approaches use metrics to measure the classifiers state.
Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
%
Some previous work also aimed to provide a benchmark suite for active learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
\end{equation}
where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
A good AL algorithm should be able to perform well in both scenarios. 
Since AL has high variance and is prone to outliers, we propose to report median AUC instead of mean AUC values. \\ [1mm]
The AUC is dependent on the chosen budget, so we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts}
	\caption{Random draws from an experimental distribution on the Splice dataset with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repetitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.
Fixing for example the classifier initialization to random draws from a constant seed does not have a corresponding use-case in the real world.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Oracle}\label{sec:oracle}
Posing active learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure would be the set that induces the highest AUC score for a given budget.
However, since this problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence. 
\cite{zhou2021towards} has used simulated annealing to search for the optimal sequence and used the best  solution found after a fixed time budget. 
Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
Our oracle simply evaluates every data point $u_k = \operatorname{unif(\mathcal{U} | k = 1 \ldots \tau)}$ in a subsample of unlabeled points by fitting the classifier on $\mathcal{L}^{(i)} \cup u_k$ and directly measuring the resulting test performance.
The point with the best test performance is selected and added to the labeled pool for that iteration.
We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
This can happen for example, if the oracle picked a labeled set that enables the classifier to correctly classifier a big portion of easy samples in the test set, but now fails to find the next single unlabeled point that would enable the classifier to succeed on one of the hard samples in the test set.
This leads to a situation, where the selected data point is basically random.\\
\begin{minipage}{0.5\linewidth}

	To circumvent this problem, we introduced margin sampling \cite{wang2014new} as a fallback option for the oracle.
	Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
	The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
	In the algorithm $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
	$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
	When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
	Alg. \ref{alg:oracle} replaces the algorithm in the AL process.
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.45\linewidth}
	\begin{algorithm}[H]
		\caption{Oracle}\label{alg:oracle}
		\begin{algorithmic}[1]
			\Require $\mathcal{U}, \mathcal{L}, \mathcal{Y}, \mathcal{D}_\text{test}$ Train, Margin, $\tau, \hat y_\theta$ 
			\State $\text{acc} \gets \operatorname{Train}(\mathcal{L}, \mathcal{D}_\text{test}, \hat y_\theta)$ 
			%\State $u \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}|)$
			\State $r^* \gets 0$
			\For{$t := 1 \ldots \tau$}
			\State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(\mathcal{U}_t, y_t)\}$
			\State $\text{acc}' \gets \operatorname{Train}(\mathcal{L}', \mathcal{D}_\text{test}, \hat y_\theta)$  
			\State $r \gets \text{acc} - \text{acc}'$
			\If{$r > r^*$} 
			\State $r^* \gets r$
			\State $u^* \gets \mathcal{U}_t$
			\EndIf
			\EndFor
			\If{$r^* = 0$}
			\State $u^* \gets \operatorname{margin}(\mathcal{U}, \hat y_\theta)$
			\EndIf
			\Return $u^*$
		\end{algorithmic}
	\end{algorithm}
\end{minipage}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Available Information}
At each iteration $i$ the AL algorithm needs to pick an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \theta^{(i)}, \text{opt}_\theta\}$, where $\theta^{(i)}$ is the current classifier and $\text{opt}_\theta$ is the optimizer used to fit $\theta^{(i)}$.
We allow algorithms to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for the classifier, or the test/validation set.

@@

%\begin{table}[]
%	\centering
%	\begin{tabular}{l|lll}
%		& Splice        & DNA           & USPS          \\
%		\hline
%		Oracle          & 0.830 +- 0.01 & 0.836 +- 0.02 & 0.823 +- 0.01 \\
%		SAL & 0.799 +- 0.01 & 0.797 +- 0.03 & 0.809 +- 0.01 \\
%		Coreset & 0.800 +- 0.01 & 0.795 +- 0.03 & 0.787 +- 0.02 \\
%		TypiClust       & 0.790 +- 0.01 & 0.771 +- 0.04 & 0.761 +- 0.02 \\
%		MarginScore     & 0.797 +- 0.02 & 0.795 +- 0.04 & 0.808 +- 0.01 \\
%		ShannonEntropy  & 0.799 +- 0.02 & 0.794 +- 0.04 & 0.807 +- 0.01 \\
%		RandomAgent     & 0.788 +- 0.01 & 0.765 +- 0.03 & 0.772 +- 0.01 \\
%		Badge           & 0.807 +- 0.01 & 0.769 +- 0.06 & 0.797 +- 0.02 \\
%		BALD            & 0.811 +- 0.01 & 0.743 +- 0.04 & 0.717 +- 0.05
%	\end{tabular}
%	\caption{Median AUC values and standard deviation for all algorithms on the tabular datasets. Higher is better.}
%\end{table}
%
\begin{figure}
	\centering
	\includegraphics[width=0.49\linewidth]{img/eval_splice}
	\hspace{1mm}
	\includegraphics[width=0.49\linewidth]{img/eval_dna}\\[2mm]
	\includegraphics[width=0.49\linewidth]{img/micro_splice}
	\hspace{1mm}
	\includegraphics[width=0.49\linewidth]{img/micro_dna}
	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
	\label{fig:eval_vector}
\end{figure}
%
\begin{figure}
\centering
	\includegraphics[width=0.49\linewidth]{img/macro_vector}
	\includegraphics[width=0.49\linewidth]{img/macro_vector_enc}
	\includegraphics[width=0.49\linewidth]{img/macro_img}
	\includegraphics[width=0.49\linewidth]{img/macro_img_enc}
	\includegraphics[width=0.49\linewidth]{img/macro_text}
	\includegraphics[width=0.49\linewidth]{img/macro_all.jpg}
	\caption{Critical Difference Diagram for all algorithms grouped by domain and all domains combined. Ranks are computed based on median AUC for each algorithm and dataset combination. Lower ranks are better.}
	\label{fig:cd_diagrams}
\end{figure}
%
From Fig. \ref{fig:eval_vector} we notice drastically different qualities for the same AL algorithm for different datasets.
We would like to highlight that both datasets are tabular from the medical domain with similar number of features and classes, yet we see that i.e. BALD is the best algorithm for Splice and the worst algorithm for DNA.
These inconsistencies are present between the datasets of all our tested domains, further highlighting the difficulties for comparing AL algorithms in terms of average performance.
In order to provide a meaningful analysis of which algorithm can be expected to perform best on average we ranked the algorithm for each dataset based on their median AUC and displayed these rankings in critical difference diagrams \cite{IsmailFawaz2018deep}.
In Fig. \ref{fig:cd_diagrams} we report the rankings split by domain as well as across all domains (excluding the toydata). \\ [2mm]
%
BALD performs bad with linear classifiers since they are trained without dropout and cannot cope well with missing inputs. \\
TypiClust is better with embedded data not only due to lower budgets. On other datasets it is not able to outperform other algorithms in early stages

\section{Conclusion}
Domains are very different \\
Even within one domain we have stark differences (Fig. \ref{fig:eval_vector}) \\
Best 3 ranks for all domains are 3, 3.1 and 3.4 ... No clear winner \\
At least on average everything is better that random

@@

\section{Alternative Plot for Restarts Ablation}
\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts_2}
\end{figure}
